---
title: "Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics"
author: "Owen Petchey, Rainer M Krug"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### import libraries and source files

```{r library_sourve, message=FALSE, include=FALSE}

# library import

library(tidyverse)
library(microxanox) # version 0.9.2
library(ggplot2)

```

### set up parameterset

A wait time were set short to be able to simulate many simulation steps and not having long simulation durations at the same time. Short wait times mean fast environmental change.
Number of strains stays one, means no investigations in trait variation in this work.
```{r param}

wait_time <- 10 # time spent at each step
num_strains <- 1 # number of strains per group
event_interval <- 10

## oxygen diffusivity equal to sulfate diffusivity means perfect symmetry
## oxygen diffusivity time series: rep(8e-2, 4), , 10 ^ c(rep(-8, 5),rep(0,5),rep(-8, 5))

# a_S at 0.125

log10a_series <- seq(3, -5, length = 10) # symmetry axis at -1, +- 4 units in both direction


num_CB_strains <- num_strains
num_SB_strains <- num_strains
num_PB_strains <- num_strains

sp <- new_strain_parameter(
  n_CB = num_CB_strains,
  values_CB = "symmetric",
  n_PB = num_SB_strains,
  values_PB = "symmetric",
  n_SB = num_PB_strains,
  values_SB = "symmetric",
  values_other  = "symmetric",
  values_initial_state = "symmetric"
  
)

parameter <- new_runsim_parameter(
  dynamic_model = bushplus_dynamic_model,  
  event_definition = event_definition_2,       
  event_interval = event_interval,
  noise_sigma = 0,
  minimum_abundances = c(1, 0 , 1),       # PB stays 0
  strain_parameter = sp,
  log10a_series =  log10a_series
)

names(parameter$minimum_abundances) <- c("CB", "PB", "SB")
rm(sp)
parameter$sim_duration <- wait_time * length(parameter$log10a_series)
parameter$sim_sample_interval <- wait_time  # to avoid having negitive ODE results

```


### scenarios from result sections

Simulation duration in each simulation is different to emphasize important dynamics in plots
To investigate in alternative stable states the CB = SB case as well as the differing in initial abundance and favouring a_O scenarios were simulated as described in the report.
To investigate in bistability the bistabilty scenario has been created, as well as adding gradual environmental change is used to induce catatrophic shifts.
All scenario dynamics are saved as RDS files.

```{r expl}
## simulate scenarios

parameter$strain_parameter$a_S <- 0.1

# case CB = SB (no events)

length <- 1e2

parameter$event_interval <- 1e6

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_unstable_eq_noevent.RDS")

# case CB = SB (events)

length <- 6e2

parameter$event_definition <- event_definition_2
parameter$event_interval <- 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_unstable_eq_event.RDS")

# different initial abundance

length <- 4e2

parameter$strain_parameter$initial_state["CB_1"] <- 11
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_different_initial_abundance_CB.RDS")

length <- 4e2

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 11

parameter$log10a_series <- rep(log10(0.1), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_different_initial_abundance_SB.RDS")

# favouring oxygen div

length <- 1.5e2

parameter$strain_parameter$initial_state["CB_1"] <- 1e5
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- rep(log10(0.001), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_favouring_a_O_SB.RDS")

length <- 1.5e2

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 1e5

parameter$log10a_series <- rep(log10(10), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_favouring_a_O_CB.RDS")

# bistability

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 1e5
parameter$log10a_series <- rep(log10(0.15), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_bistabilty_SB.RDS")

parameter$strain_parameter$initial_state["CB_1"] <- 1e5 
parameter$strain_parameter$initial_state["SB_1"] <- 10
parameter$log10a_series <- rep(log10(0.15), times = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_bistabilty_CB.RDS")

# gradual environmental change to inestigate in catastrophic shifts

length <- 1e3
times <- length / 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- c(rep(-5, times = times), seq(-5, 3, length = length), rep(3, times = times), seq(3, -5, length = length), rep(-5, times = times))
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_varying_a_O_both_collapse.RDS")

length <- 1e3
times <- length / 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- c(rep(-5, times = times), seq(-5, 3, length = length), rep(3, times = times))
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_varying_a_O_SB_collapse.RDS")

length <- 1e3
times <- length / 10

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10

parameter$log10a_series <- c(rep(3, times = times), seq(3, -5, length = length), rep(-5, times = times))
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

results <- run_simulation(parameter)
saveRDS(results, file = "RDS_files/sym_dynamics_varying_a_O_CB_collapse.RDS")

```

### plotting scenarios

Scenarios described above are plotted with the plot_dynamics function defined in functions.R after read in from RDS file.

```{r plot_1, include=FALSE}
## ggplot

eq_noevent_results <- readRDS(file = "RDS_files/sym_dynamics_unstable_eq_noevent.RDS")
p1.1 <- plot_dynamics_symmetric(eq_noevent_results, plot_a = FALSE)

eq_event_results <- readRDS(file = "RDS_files/sym_dynamics_unstable_eq_event.RDS")
p1.2 <- plot_dynamics_symmetric(eq_event_results, plot_a = FALSE)

init_ab_CB_results <- readRDS(file = "RDS_files/sym_dynamics_different_initial_abundance_CB.RDS")
p2.1 <- plot_dynamics_symmetric(init_ab_CB_results, plot_a = FALSE)

init_ab_SB_results <- readRDS(file = "RDS_files/sym_dynamics_different_initial_abundance_SB.RDS")
p2.2 <- plot_dynamics_symmetric(init_ab_SB_results, plot_a = FALSE)

favouring_CB_results <- readRDS(file = "RDS_files/sym_dynamics_favouring_a_O_CB.RDS")
p3.1 <- plot_dynamics_symmetric(favouring_CB_results, plot_a = FALSE)

favouring_SB_results <- readRDS(file = "RDS_files/sym_dynamics_favouring_a_O_SB.RDS")
p3.2 <- plot_dynamics_symmetric(favouring_SB_results, plot_a = FALSE)

bistability_CB_results <- readRDS(file = "RDS_files/sym_dynamics_bistabilty_CB.RDS")
p4.1 <- plot_dynamics_symmetric(bistability_CB_results, plot_a = FALSE)

bistability_SB_results <- readRDS(file = "RDS_files/sym_dynamics_bistabilty_SB.RDS")
p4.2 <- plot_dynamics_symmetric(bistability_SB_results, plot_a = FALSE)

a_O_vary_SB_results <- readRDS(file = "RDS_files/sym_dynamics_varying_a_O_SB_collapse.RDS")
p5.1 <- plot_dynamics_symmetric(a_O_vary_SB_results, plot_a = TRUE)

a_O_vary_CB_results <- readRDS(file = "RDS_files/sym_dynamics_varying_a_O_CB_collapse.RDS")
p5.2 <- plot_dynamics_symmetric(a_O_vary_CB_results, plot_a = TRUE)

a_O_vary_both_results <- readRDS(file = "RDS_files/sym_dynamics_varying_a_O_both_collapse.RDS") # add collapse next time run sim above
p5.3 <- plot_dynamics_symmetric(a_O_vary_both_results, plot_a = TRUE)

# chnagel labels smaller or bigger fizre


# discussion pssouble further ideas
```

### saving plots

Plots are saved with ggplot

```{r ggsave_1}
width = 30
height = 15
units = "cm"
## p1
ggsave("GG_figures/eq_noevent_figure.pdf", plot = p1.1, width = width, height = height, units = units)
ggsave("GG_figures/eq_event_figure.pdf", plot = p1.2, width = width, height = height, units = units)
## p2
ggsave("GG_figures/diff_init_ab_CB_figure.pdf", plot = p2.1, width = width, height = height, units = units)
ggsave("GG_figures/diff_init_ab_SB_figure.pdf", plot = p2.2, width = width, height = height, units = units)
## p3
ggsave("GG_figures/favouring_a_O_CB_figure.pdf", plot = p3.1, width = width, height = height, units = units)
ggsave("GG_figures/favouring_a_O_SB_figure.pdf", plot = p3.2, width = width, height = height, units = units)
## p4
ggsave("GG_figures/bistabilty_CB_figure.pdf", plot = p4.1, width = width, height = height, units = units)
ggsave("GG_figures/bistabilty_SB_figure.pdf", plot = p4.2, width = width, height = height, units = units)
## p5
ggsave("GG_figures/a_O_gradual_vary_SB_collapse_figure.pdf", plot = p5.1, width = width, height = height, units = units)
ggsave("GG_figures/a_O_gradual_vary_CB_collapse_figure.pdf", plot = p5.2, width = width, height = height, units = units)
ggsave("GG_figures/a_O_gradual_vary__collapse_figure.pdf", plot = p5.3, width = width, height = height, units = units)

rm(width, height, units)
```

Display plots in file:

```{r plot_2}
p1.1
p1.2
p2.1
p2.2
p3.1
p3.2
p4.1
p4.2
p5.1
p5.2
p5.3

# split ylab \n
```

### temporal results

Gradual envrionmental change is simulated in both direction with the temporal simulation method. Two simulation durations of 1000 respectively 10'000 hours to emphasize problem of bistability with slower environmental change. Simulations saved as RDS files.

```{r temporal_2}
length <- 1e4 # modify sim duration

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10
parameter$log10a_series <- log10a_series <- seq(-5, 3, length = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

temporal_results <- run_temporal_ssfind(parameter) ## simulate
saveRDS(temporal_results, file = "RDS_files/temporal_ss_10000h.RDS")

length <- 1e3 # modify sim duration

parameter$strain_parameter$initial_state["CB_1"] <- 10
parameter$strain_parameter$initial_state["SB_1"] <- 10
parameter$log10a_series <- log10a_series <- seq(-5, 3, length = length)
parameter$sim_duration <- length(parameter$log10a_series) * wait_time 

temporal_results <- run_temporal_ssfind(parameter) ## simulate
saveRDS(temporal_results, file = "RDS_files/temporal_ss_1000h.RDS")
```

Plotting of temporal simulations with plot_temporal_ss from functions.R

```{r plot_3, include=FALSE}
temporal_results_1000h <- readRDS(file = "RDS_files/temporal_ss_1000h.RDS")
p6.1 <- plot_temporal_ss(temporal_results_1000h, zoom = FALSE) ## plot

temporal_results_10000h <- readRDS(file = "RDS_files/temporal_ss_10000h.RDS") # slower environmental change
p6.2 <- plot_temporal_ss(temporal_results_10000h, zoom = FALSE) ## plot

# H: because only a_O chnages
# discusion: vary a_O & a_S, longer simulations, optimazation of symmetry 
# save as GG
```

Displaying plots

```{r plot_4}
p6.1
p6.2

```
GGsave the plots as above if required

```{r ggsave_4}
height = 30
width = 15
units = "cm"

ggsave("GG_figures/temporal_1000h_figure.pdf", plot = p6.1, width = width, height = height, units = units)

ggsave("GG_figures/temporal_10000h_figure.pdf", plot = p6.2, width = width, height = height, units = units)

```
